---
title: "Section 6 Exercises"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Section 6.6

```{r}
library(tidyverse)
hpi_df <- read_csv("data/hpi-tidy.csv")
```

### Exercise 1

```{r}
world_df <- ggplot2::map_data("world")
world_full <- left_join(world_df, hpi_df, by = c("region" = "Country"))

ggplot(data = world_full, aes(x = long, y = lat, group = group)) + 
  geom_polygon(colour = "black", aes(fill = HappyPlanetIndex)) + 
  coord_map(projection = "mercator") +
  theme_void() +
  scale_fill_viridis_b()
```

### Exercise 2

```{r}
world_df <- world_df %>% mutate(region = if_else(region == "USA", "United States of America", region))
world_full <- left_join(world_df, hpi_df, by = c("region" = "Country"))

ggplot(data = world_full, aes(x = long, y = lat, group = group)) + 
  geom_polygon(colour = "black", aes(fill = HappyPlanetIndex)) + 
  coord_map(projection = "mercator") +
  theme_void() +
  scale_fill_viridis_b()
```

### Exercise 3

```{r}
ggplot(data = world_full, aes(x = long, y = lat, group = group)) + 
  geom_polygon(colour = "black", aes(fill = HappyPlanetIndex)) + 
  coord_map(projection = "mercator", xlim=c(-180,180)) +
  theme_void() +
  scale_fill_viridis_b()
```

### Exercise 4

The blog post explains that the Mercator projection distorts sizes of countries further from the equator. An example is Greenland: while its true size could fit into Africa 14 times, this projection makes it look the size of, if not larger than Africa.

### Exercise 5

```{r}
ggplot(data = world_full, aes(x = long, y = lat, group = group)) + 
  geom_polygon(colour = "black", aes(fill = HappyPlanetIndex)) + 
  coord_map(projection = "globular", xlim=c(-180,180)) +
  theme_void() +
  scale_fill_viridis_b()

ggplot(data = world_full, aes(x = long, y = lat, group = group)) + 
  geom_polygon(colour = "black", aes(fill = HappyPlanetIndex)) + 
  coord_map(projection = "gilbert", xlim=c(-180,180)) +
  theme_void() +
  scale_fill_viridis_b()
```

Globular seems to keep Greenland the correct size, while gilbert makes the map round but Greenland stays much larger than it should be.