---
title: 'Section 2: Grammar of Graphics'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(palmerpenguins)
library(tidyverse)
penguins

## single quantitative frequency plot
ggplot(data = penguins, aes(x = bill_length_mm)) +
  geom_freqpoly()
## single categorical barplot
ggplot(data = penguins, aes(x = species)) +
  geom_bar()
## two quant scatterplot
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  geom_smooth()
## two cat heatmap
ggplot(data = penguins, aes(x = species, y = island)) +
  geom_bin_2d()
ggplot(data = penguins, aes(x = bill_length_mm)) +
  geom_histogram(colour = "black", fill = "white") +
  facet_wrap(~ species, ncol = 1)
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  geom_smooth() +
  facet_wrap(~ species)
```
```{r}
ggplot(data = penguins, aes(x = species)) +
  geom_bar(stat = "count") ## don't need stat = "count" because it's the default.

penguins_count <- penguins %>% group_by(species) %>% 
  summarise(ncount = n())

## graphs bar plot when variable has count
ggplot(data = penguins_count, aes(x = species, y = ncount)) +
  geom_bar(stat = "identity")

ggplot(data = penguins, aes(x = species, fill = sex)) +
  geom_bar(position = "stack") + ## don't need "stacked" because it's default
  scale_fill_viridis_d()

penguins_nona <- penguins %>% filter(!is.na(sex))

## side-by-side bar plot
ggplot(data = penguins_nona, aes(x = species, fill = sex)) +
  geom_bar(position = "dodge") + ## don't need "stacked" because it's default
  scale_fill_viridis_d()
ggplot(data = penguins, aes(x = species)) +
  geom_bar() +
  coord_polar()
ggplot(data = penguins, aes(x = species)) + 
  geom_bar() +
  coord_flip()
```

Exercise: Make a plot of the penguins data that explicitly specifies all 7 parameters in the grammar of graphics, even if you just explicitly specify the default for some of the parameters.

```{r}
ggplot(data = penguins, (aes(x = island, fill = sex))) + 
  geom_bar(stat = "count", position = "dodge") + facet_wrap(~ species) + coord_flip()
```

```{r}
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm,
                            colour = species)) +
  geom_point() +
  geom_smooth()
```

```{r}
library(tidyverse)
category <- rep(c("Group 1", "Group 2", "Group 3", "Group 4", "Group 5"), 7)
time <- rep(c(2000, 2001, 2002, 2003, 2004, 2005, 2006), each = 5) 
response <- rpois(length(category), 30)
df <- tibble(category, time, response)
ggplot(data = df, aes(x = time, y = response)) +
  geom_col(aes(fill = category)) +
  labs(y = "count") +
  scale_fill_viridis_d()

## position changed
ggplot(data = df, aes(x = time, y = response)) +
  geom_col(aes(fill = category), position = "dodge") +
  labs(y = "count") +
  scale_fill_viridis_d()

## faceted bar plot
ggplot(data = df, aes(x = time, y = response)) +
  geom_col(aes(fill = category)) +
  labs(y = "count") +
  scale_fill_viridis_d() + 
  facet_wrap(~ category)

## line plot
ggplot(data = df, aes(x = time, y = response, colour = category)) + 
  geom_line() +
  scale_colour_viridis_d()
```

## Exercise 1.
In the above plots, we skipped the final “two variable” combination: one quantitative variable and one categorical variable. Create either a side-by-side boxplot or side-by-side violin plot of a quantitative and a categorical variable in the penguins data set.

```{r}
ggplot(data = penguins, aes(x = sex, y = body_mass_g)) + geom_boxplot()
```

## Exercise 2.
Modify the frequency plot made with geom_freqpoly() to use colour in two different ways:

change the colour of the line to any colour that R knows. A list of some colours can be found at this link.

add a colour asethetic to make three different frequency lines, one for each species of penguin.

Hint: recall that only aesthetic mappings (variables) go inside aes(). Of (a) or (b), which one is an aesthetic mapping?

```{r}
ggplot(data = penguins, aes(x = bill_length_mm)) +
  geom_freqpoly(colour = "blue")

ggplot(data = penguins, aes(x = bill_length_mm, colour = species)) +
  geom_freqpoly()
```

## Exercise 3.
Thus far, we have only plotted one or two variables at a time. Recall that one way to construct a plot of three different variables is to use colour. Modify the scatterplot so that the points are coloured by species.

```{r}
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm, colour = species)) +
  geom_point() +
  geom_smooth()
```

## Exercise 4.
Recall that aes() aesthetics specified within the ggplot() function directly are called global aesthetics because every other GEOM will use them (unless specifically overridden) while aes() specified within a particular GEOM are called local aesthetics because only that particular GEOM will use them.

Modify the scatterplot and smoother so that

the points are coloured by species, but there is only one smoother instead of three.

there are three different coloured smoothers (one for each species) but the points are all the same colour.

```{r}
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(colour = species)) +
  geom_smooth()

library(plotly)

g <- ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm, label = island)) +
  geom_point() +
  geom_smooth(aes(colour = species))

g

ggplotly(g, tooltip = "label")
```

## Section 2.3

```{r}
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm,
                            colour = species)) +
  geom_point() +
  geom_smooth() + 
  theme(plot.background = element_rect(fill = "yellow"), axis.ticks.length = unit(1.1, "cm"), legend.key = element_rect(fill = "darkorange3", colour = "blue"), legend.text = element_text(size = 8, colour = "red"), axis.title.y = element_text(size = rel(1.5), angle = 45), legend.position = c(0,0), legend.direction = "horizontal", aspect.ratio = 5)
```