---
title: '`tidyverse` Review'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Aside: Baseline of Zero?

```{r}
library(tidyverse)
library(palmerpenguins)
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) + 
  geom_point() + 
  xlim(c(0, 60)) + 
  ylim(c(0, 25))
## length vs relative position: people see bar plots as being comparisons of length, while a line plot or scatterplot can have a different scale because people can conceptualize the relative difference
```

## Section 5.1 Exercises: Billboard

The data set only contains a few variables:

- `no`, the number on the billboard year end Top 100 (1 - 100)
- `title`, the title of the song
- `artist`, the artist of the song
- `year`, the year of the song

```{r}
library(billboard)
head(wiki_hot_100s)
tail(wiki_hot_100s)
max(wiki_hot_100s$year)
library(rvest)
library(tidyverse)

billboard_count <- wiki_hot_100s %>% filter(year >=2000 & year < 2010) %>% 
  group_by(artist) %>% 
  summarise(ncount = n()) %>% 
  arrange(desc(ncount)) %>% 
  slice(1:10) %>% 
  mutate(artist_ordered = fct_reorder(.f = artist, .x = ncount))

ggplot(data = billboard_count, aes(x = ncount, y = artist_ordered, fill = artist_ordered)) + 
  geom_col() + 
  scale_fill_viridis_d()
```

## Exercises

Exercise 1. Make the visualization that we sketched in class. We will complete this exercise as a class. (done)

Exercise 2. There is a minor flaw in the way that we counted up the number of hits for each artist. Examine the 2nd to last row of the original data set with tail() to look at this potential flaw. What do you find?

Any song that features an artist will not count toward their total top 100 songs.

Exercise 3. Challenging. Fix the issue in Exercise 2. May want to skip this question and come back after completing the other exercises.

```{r}
sep_artist <- wiki_hot_100s %>% separate(col = artist, into = c("artist1", "artist2"), sep = " featuring ") %>% 
  pivot_longer(cols = c(artist1, artist2), names_to = "name", values_to = "artist") %>% 
  filter(!is.na(artist)) %>% 
  select(-name)
```

Exercise 4. Change the plot from Exercise 1 to be a Lollipop chart using this website as a reference. Why might the lollipop chart be better than a bar plot?

```{r}
ggplot(data = billboard_count, aes(x = ncount, y = artist_ordered)) + 
  geom_point() + 
  geom_segment(aes(x = ncount, xend = ncount, y = 0, yend = artist_ordered))
```

With a lollipop chart, it is easier to compare which artists had the same number of top 100 hits over the 10 year period. Higher data-to-ink ratio.

Exercise 5. Use this website to customize the end points of your lollipop chart. If you have time, you can explore the other customization options. Make it look fancy!

```{r}
ggplot(data = billboard_count, aes(x = ncount, y = artist_ordered)) + 
  geom_point(colour="orange", size=4) + 
  geom_segment(aes(x = ncount, xend = ncount, y = 0, yend = artist_ordered), colour = "grey")
```

```{r}
## provide the URL and name it something (in this case, url).
## paste0 pastes together the base URL and the year into a single string:
## this will be useful in a moment
year <- 2017

url <- paste0("https://en.wikipedia.org/wiki/Billboard_Year-End_Hot_100_singles_of_", year)
h <- url %>% 
  httr::GET(config = httr::config(ssl_verifypeer = FALSE)) %>% 
  read_html()  

## grabs the tables
tab <- h %>% html_nodes("table")
df <- tab[[1]] %>% html_table() %>%
  mutate(year = 2017)
df
get_wiki_100 <- function(year) {
  
  ## same code as before, replacing 2017 with year.
url <- paste0("https://en.wikipedia.org/wiki/Billboard_Year-End_Hot_100_singles_of_", year)
h <- url %>% 
  httr::GET(config = httr::config(ssl_verifypeer = FALSE)) %>% 
  read_html() 
  
  tab <- h %>% html_nodes("table")
  df <- tab[[1]] %>% html_table() %>%
    mutate(year = year)
  
  ## tell our function to return the dataframe `df`
  return(df) 
}
get_wiki_100(year = 2017)
library(purrr)
year_list <- list(2017, 2018, 2019, 2020, 2021)
year_list

df_all <- map(year_list, get_wiki_100)
df_all ## a list of data frames, one for each year
df_2017_present <- bind_rows(df_all)
df_2017_present <- df_2017_present %>%
  mutate(Title = str_remove_all(Title, pattern = "\"")) %>% ## get rid of \ in title
  rename(no = No., 
         title = Title, 
         artist = `Artist(s)`) ## make column names match with billboard package

wiki_tibble <- as_tibble(wiki_hot_100s) %>% ## convert billboard data to tibble
  mutate(year = as.numeric(year),
         no = as.integer(no)) ## change variable types to match with scraped data

hot100_df <- bind_rows(wiki_tibble, df_2017_present)
```

Exercise 6. Use the hot100_df to make either a bar plot or a lollipop chart of the most popular artists of the 2010s (2010 through 2019). It may be helpful to make this plot without looking back at the code for the 2000s plot until you get stuck.

```{r}
top10_df <- hot100_df %>% filter(year <=2019 & year >= 2010) %>% 
  group_by(artist) %>% 
  summarise(nsongs = n()) %>% 
  arrange(desc(nsongs)) %>% 
  slice(1:10) %>% 
  mutate(artist = fct_reorder(artist, nsongs))

ggplot(data = top10_df, aes(x = nsongs, y = artist, label = artist)) + 
  geom_point() + 
  geom_segment(aes(x = 0, xend = nsongs, y = artist, yend = artist))
```

Exercise 7. Much of the code to scrape the data, using purrr to iterate over the scrape, and then combining the list of data frames to a single data frame may be new. It is not expected that you are able to write this code on your own, but you should have an overall understanding of what the code is doing. Write 2-3 sentences that summarizes the overall purpose of the rvest and purrr code.

The rvest package scrapes data from the Wikipedia website while the purrr package helps us to use our function over and over instead of typing out every year we want in order to get a full data set with all the info we want. We then use ggplot to analyze the data collected.

Exercise 8. If you have time, use purrr and rvest to read different data from Wikipedia into R. For example, you might consider sports championship data, which will typically have a different Wikipedia page for each year or season.



## Section 5.2: HPI

```{r}
library(tidyverse)
hpi_df <- read_csv("data/hpi-tidy.csv")
hpi_df
ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing)) +
  geom_point()
hpi_us <- hpi_df %>% filter(Country == "United States of America")
hpi_us
ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing)) +
  geom_point() +
  geom_label(data = hpi_us, aes(label = Country)) ## specify
## data = hpi_us so geom_label only uses the observation in hpi_us
library(ggrepel)
ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing)) +
  geom_point() +
  geom_label_repel(data = hpi_us, aes(label = Country)) +
  geom_point(data = hpi_us, size = 3, shape = 1) ## create a second point that is an open circle (shape = 1) with a larger size (size = 3) to surround the United States point on the scatterplot
```

Exercise 1. Change the code to label 3 countries of interest. Recall that you will need to use the | operator in the dplyr::filter() function.

```{r}
hpi_interest <- hpi_df %>% filter(Country == "Afghanistan" | Country == "Albania" | Country == "Algeria")

ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing)) +
  geom_point() +
  geom_label_repel(data = hpi_interest, aes(label = Country)) +
  geom_point(data = hpi_interest, size = 3, shape = 1)
```

### `plotly`

```{r}
## install.packages("plotly")
library(plotly)
plot1 <- ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing)) +
  geom_point()
ggplotly(plot1)
plot1 <- ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing,
                                   label = Country)) +
  geom_point()
ggplotly(plot1, tooltip = "label")
```

Exercise 2. Use the ggplotly() function on any other plot we’ve made so far.

In `02-grammar-of-graphics.rmd` file

Exercise 3. What are some advantages of making a plot more interactive with plotly? What are some disadvantages?

advantages: interesting, allows user to explore, good way to present lots of data
disadvantages: requires a computer, adds extra "stuff", difficult to compare multiple points, hard to make a particular point (western countries have higher footprints)

```{r}
ggplot(data = hpi_df, aes(x = Footprint, y = Wellbeing)) +
  geom_point() +
  geom_label_repel(data = hpi_us, aes(label = Country)) +
  geom_point(data = hpi_us, size = 3, shape = 1) +
  labs(title = "Countries with a Higher Ecological Footprint Tend to Have Citizens with Higher Wellbeing", ## add title
    subtitle = "Wellbeing is on a 1-10 scale", ## add subtitle (smaller text size than the title)
    caption = "Data Source: http://happyplanetindex.org/countries", ## add caption to the bottom of the figure
    x = "Ecological Footprint", ## change x axis label
    y = "Wellbeing") ## change y axis label
```

Exercise 4. In which of those 3 situations are we in for the World Region graph if we were to colour by Region?

unordered

```{r}
ggplot(data = hpi_df, aes(x = Footprint, y = HappyLifeYears, colour = Region)) +
  geom_point() +
  scale_colour_brewer(palette = "Accent")

ggplot(data = hpi_df, aes(x = Footprint, y = HappyLifeYears, colour = Region)) +
  geom_point() +
  scale_colour_brewer(palette = "Dark2")

ggplot(data = hpi_df, aes(x = Footprint, y = HappyLifeYears, colour = Region)) +
  geom_point() +
  scale_colour_viridis_d(option = "plasma")
```

Exercise 6. Read the examples section of the Help file for ?scale_colour_viridis_d. What’s the difference between scale_colour_viridis_d(), ?scale_colour_viridis_c(), and scale_colour_viridis_b()?

d: discrete, c: continuous, b: bin continuous variable

```{r}
ggplot(data = hpi_df, aes(x = Footprint, y = HappyLifeYears)) +
  geom_point() +
  facet_wrap( ~ Region)
library(palmerpenguins)
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point(aes(colour = species)) ## colour is good enough here
ggplot(data = penguins, aes(x = bill_length_mm, y = bill_depth_mm)) +
  geom_point() +
  facet_wrap( ~ species) ## faceting probably unnecessary: colour is better

## faceting is more useful if:
## - there are "more" categories AND
## - there is a lot of overlap between groups

colour_bad <- tibble(x = rnorm(500, 0, 1), y = rnorm(500, 0, 1),
       groupvar = c(rep("A", 50), rep("B", 50),
                    rep("C", 50), rep("D", 50),
                    rep("E", 50), rep("F", 50), rep("G", 50),
                    rep("H", 50), rep("I", 50), rep("J", 50)))

ggplot(data = colour_bad, aes(x = x, y = y, colour = groupvar)) +
  geom_point() + ## can't distinguish anything really: colour is bad
  geom_smooth(se = FALSE)
ggplot(data = colour_bad, aes(x = x, y = y)) +
  geom_point() +
  geom_smooth(se = FALSE) +
  facet_wrap( ~ groupvar) ## faceting better
```

