---
title: "Quiz 6 In Class"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

“All work presented is my own. I have not communicated with or worked with anyone else on this quiz.” - Maddox Ormiston

## Question 1

Using the majors data, one might want to visualize the breakdown of gender for each major. To do this, one would construct bar plots faceted by gender, and reordered by count, to see the difference in the number of people of both genders who major in different disciplines. A static graph would be better because we are only trying to visualize one "thing", and interaction is unnecessary.

## Question 2

```{r}
library(shiny)
library(tidyverse)
library(here)
atp_df <- read_csv(here("data/atp_matches_2019.csv"))
wta_df <- read_csv(here("data/wta_matches_2019.csv"))
both_df <- bind_rows(atp_df, wta_df)

both_long <- both_df %>% pivot_longer(c(winner_name, loser_name))

## only keep players who have player over 50 matches
both_n50 <- both_long %>% group_by(value) %>% count() %>%
  filter(n > 50)

## construct various statistics
major_tennis <- semi_join(both_long, both_n50, by = c("value"))
major_tennis <- major_tennis %>% mutate(w_svperc = 100 * w_1stIn / w_svpt,
                        l_svperc = 100 * l_1stIn / l_svpt,
                        w_firstwon = 100 * w_1stWon / w_1stIn,
                        l_firstwon = 100 * l_1stWon / l_1stIn,
                        w_secondwon = 100 * w_2ndWon / (w_svpt - w_1stIn),
                        l_secondwon = 100 * l_2ndWon / (l_svpt - l_1stIn))

major_tennis_w <- major_tennis %>% filter(name == "winner_name")
major_tennis_l <- major_tennis %>% filter(name == "loser_name")

w_small <- major_tennis_w %>% select(value, winner_seed, w_ace, w_df, w_svperc,
                                     w_firstwon, w_secondwon) %>%
  rename(seed = winner_seed, ace = w_ace, df = w_df, svperc = w_svperc,
         firstwon = w_firstwon, secondwon = w_secondwon)

l_small <- major_tennis_l %>% select(value, loser_seed, l_ace, l_df, l_svperc, l_firstwon, l_secondwon)  %>%
  rename(seed = loser_seed, ace = l_ace, df = l_df, svperc = l_svperc,
         firstwon = l_firstwon, secondwon = l_secondwon)

df <- bind_rows(w_small, l_small) %>%
  rename(player = "value")
df

var_choices <- names(df)[3:7]

library(shiny)

ui <- fluidPage(
  sidebarLayout(sidebarPanel(
    selectizeInput("playerchoice",
                   label = "Choose a Player", choices = levels(factor(df$player)),
                   selected = "Aryna Sabalenka"),
    selectizeInput("playerchoice2", 
                   label = "Choose a second player", choices = levels(factor(df$player)),
                   selected = "Daniil Medvedev"), 
    radioButtons("varchoice", label = "Choose a Statistic",
                 choices = var_choices), 
    checkboxInput("mean", label = "Display the mean?")),
    mainPanel(plotOutput("histgraph"), 
              plotOutput("boxplot"))
  )
)

server <- function(input, output, session) {
  
  df_sub <- reactive({
    df %>% filter(player == input$playerchoice)
  })
  
  df_sub2 <- reactive({
    df %>% filter(player == input$playerchoice | player == input$playerchoice2)
  })
  
  hist_plot <- reactive({
    # ggplot(df_sub(), aes_string(x = input$varchoice)) +
    # geom_histogram(colour = "black", fill = "white", bins = 15)
    ggplot(df_sub(), aes(x = .data[[input$varchoice]])) +
      geom_histogram(colour = "black", fill = "white", bins = 15)
  })
  
  output$histgraph <- renderPlot({
    hist_plot()
  })
  
  box_plot <- reactive({
    ggplot(data = df_sub2(), aes(x = player, y = .data[[input$varchoice]])) + 
      geom_boxplot()
  })
  
  output$boxplot <- renderPlot({
    box_plot()
  })
  
  output$mean <- renderPlot({
    if(input$mean){
      box_plot() + stat_summary(fun = mean, geom = "point", colour = "purple", shape = 8,
               size = 3)
    }
  })
}

shinyApp(ui, server)
```

