---
title: "Quiz 5 Take Home"
author: Maddox Ormiston
output: 
  rmdformats::readthedown:
    code_folding: show
    toc_depth: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

“All work presented is my own, and I have followed all rules for collaboration.” -Maddox Ormiston

Worked with James Wolpe on the quiz

## Question 1

Construct a graph that uses the number of Facebook friends each student has.

```{r, message = FALSE}
library(tidyverse)
statsurvey_df <- read_csv("data/stat113_survey.csv")

statsurvey_fb1 <- statsurvey_df %>% 
  filter(!is.na(Facebook)) %>% 
  separate(time_both, into = c("semester", "year"), sep = 1) %>% 
  arrange(year, desc(semester)) %>% 
  unite(col = "time_both", c(semester, year)) %>% 
  mutate(time_both = fct_inorder(time_both))

stat_facebook <- statsurvey_fb1 %>% 
  group_by(time_both) %>% 
  summarise(meanfriend = mean(Facebook), sdfriend = sd(Facebook), nfriend = n()) %>% 
  mutate(l_se = meanfriend - sdfriend / sqrt(nfriend),
         u_se = meanfriend + sdfriend / sqrt(nfriend))

ggplot(stat_facebook, aes(x = time_both, y = meanfriend)) +
  geom_point() +
  geom_errorbar(aes(ymin = l_se, ymax = u_se)) + 
  labs(x = "Semester", y = "Average Number of Facebook Friends")
```

It looks like Facebook friends have decreased over time.

## Question 2

```{r}
statsurvey_fb2 <- statsurvey_df %>% 
  separate(time_both, into = c("semester", "year"), sep = 1) %>% 
  arrange(year, desc(semester)) %>% 
  unite(col = "time_both", c(semester, year)) %>% 
  mutate(time_both = fct_inorder(time_both))

statsurvey_fb2 <- statsurvey_fb1 %>% mutate(Facebook = replace(Facebook, Facebook == 0, NA)) %>% 
  mutate(fb = if_else(is.na(Facebook), true = "No", false = "Yes")) %>% 
  group_by(time_both, fb) %>% 
  summarise(ncount = n()) %>%
  ungroup() %>%
  group_by(time_both) %>%
  mutate(ntotal = sum(ncount)) %>%
  ungroup() %>%
  filter(fb == "Yes") %>% 
  mutate(prop = ncount / ntotal)

ggplot(data = statsurvey_fb2, aes(x = time_both, y = prop)) + 
  geom_point() + 
  labs(x = "Semester", y = "Proportion of Students Who Have Facebook")
```

This graph also shows that the proportion of students who have a Facebook acount has decreased over time.